Taking Dyads Seriously

### 2017-10-29 paper/1_intro.tex
max, you are killing me.

We begin with a brief review of these dependencies and the AME model. Next, we conduct a simulation study to show how the AME approach can recover unbiased and well-calibrated regression coefficients in the presence of the dependencies that arise in dyadic data. Then, we apply this approach to five prominent studies in the international relations (IR) literature and compare results from the current state-of-the-art approach (a GLM with robust standard errors) to those obtained using the AME framework. The comparison reveals that in ignoring observational dependence, the standard approach overestimates the effect of key variables in the literature and underestimates their uncertainty. Consequently, the latent factor approach (AME) produces results that, at times, contradict those found in these studies in particular, and the broader literatures from which they are drawn. Moreover, we demonstrate the latent factor approach offers substantive insights that are often occluded by ignoring the interdependencies in most relational data that the field of IR is concerned with. Finally, we show that for each replication our network based approach provides substantively more accurate out-of-sample predictions than the models used in the original studies. Thus, the AME approach can be used by scholars in the field to not only generate substantive insights, but it also enables us to better model the data generating process behind events of interest.  Most importantly, it facilitates concentration on the relations aspect of the field of international relations.

>>>>

We begin with a brief review of these dependencies and the AME model. Next, we conduct a simulation study to show how the AME approach can recover unbiased and well-calibrated regression coefficients in the presence of the dependencies that arise in dyadic data. Then, we apply this approach to five prominent studies in the international relations (IR) literature and compare results from the current state-of-the-art approach (a GLM with robust standard errors) to those obtained using the AME framework. The comparison reveals that in ignoring observational dependence, the standard approach overestimates the effect of key variables in the literature and underestimates their uncertainty. Consequently, the latent factor approach (AME) produces results that, at times, contradict those found in these studies in particular, and the broader literatures from which they are drawn. Moreover, we demonstrate the latent factor approach offers substantive insights that are often occluded by ignoring the interdependencies in most relational data that the field of IR is concerned with. Finally, we show that for each replication our network-based approach provides substantively more accurate out-of-sample predictions than the models used in the original studies. Thus, the AME approach can be used by scholars in the field to not only generate substantive insights, but it also enables us to better model the data generating process behind events of interest.  Most importantly, it facilitates concentration on the relations aspect of the field of international relations.

### 2017-10-29 paper/2_model.tex
max, you are killing me.

For similar reasons one also needs to take into account the shared dependence between observations that share a common receiver. The bottom-left panel, illustrates that sender and receiver type dependencies can also blend together. Specifically, actors who are more likely to send ties in a network tend to also be more likely to receive them. As a result, the rows and columns in an adjacency matrix are often correlated. For example, consider that trade flows both from and to many wealthy, developed countries. The bottom-right panel, highlights a second order dependence, specifically, reciprocity. This is a dependency occurring within dyads involving the same actors whereby values of $y_{ij}$ and $y_{ji}$ are correlated. The concept of reciprocity has deep roots in the study of relations between states \citep{richardson:1960,keohane:1989}.

For most relational data, however, dependencies do not simply manifest at the nodal or dyadic level. More often we find significant evidence of higher order structures that result from dependencies between multiple groups of actors. These dependencies arise because there may be a or some set of latent attributes between actors that affects their probability of interacting with one another \citep{zinnes:1967,wasserman:faust:1994}. In Figure~\ref{fig:thirdDeps} we provide a visualization of a simulated relational dataset wherein the nodes designate actors and edges between the nodes indicate that an interaction between the two took place. To highlight third order dependence patterns, nodes with similar latent attributes are colored similarly.

@@

The visualization illustrates that the actors belonging to the same group have a higher likelihood of having an interaction with each other, whereas interactions across groups are more rare. A prominent example of a network with this type of structure was found by \citet{adamic:glance:2005}, who visualized the ways in which right and left leaning political blogs linked to one another preceding the 2004 United States Presidential Election. \citeauthor{adamic:glance:2005} find that the degree of interaction between right and left leaning blogs was minimal, and that most blogs linked to others that were politically similar. This showcases the types of higher-order dependencies that can emerge in relational data. First, the fact that interactions were determined by a shared attribute, in this case political ideology, is an example of \textit{homophily}. Homophily explains the emergence of patterns such as transitivity (``a friend of a friend is a friend'') and balance (``an enemy of a friend is an enemy''), which also have a long history in international relations. The other major type of meso-scopic feature that emerges in relational data is community structure \citep{mucha:etal:2010}, which is often formalized through the concept of stochastic equivalence \citep{anderson:etal:1992}. This concept simply refers to the idea that groups of nodes that act similarly in the network are stochastically equivalent. In the example we have laid out above each of the left leaning blogs would be considered stochastically equivalent to one another.

The presence of these dependencies in relational data points to the fact that there is a complex structure underlying the dyadic events that we observe, and that accounting for this structure is necessary if we are to adequately represent the data generating process. Of course, if one is able to specify each of the nodal, dyadic, and triadic set of attributes that influence interactions then one can be assured that the conditional independence assumption underlying standard approaches will be satisfied. However, it is rarely the case that we can do this even for TSCS data, thus we more than often make modeling decisions to account for that structure. Failing to account for the underlying structure in either TSCS or dyadic data leads to a number of well-known challenges: a) biased estimates of the effect of independent variables, b) uncalibrated confidence intervals, and c) poor predictive performance. Further by ignoring these potential interdependencies, we often ignore substantively interesting features of the problem. The study of international relations is founded on the relations among actors. Why ignore the interdependencies that led to the study of IR in the first place?

@@

To account for the dependencies that are prevalent in dyadic data, we turn to the AME model. The AME approach can be used to conduct inference on cross-sectional and longitudinal networks with binary, ordinal, or continuous linkages. It is flexible and easy to use for analyzing the kind of relational data often found in social science. It accounts for nodal and dyadic dependence patterns, as well as higher-order dependencies such as homophily and stochastic equivalence. \citet{hoff:2015:arxiv,minhas:etal:2016:arxiv} provide a detailed introduction to this framework, and the latter piece also details how it contrasts with alternative network based approaches. Here we provide a brief review and then move to a simulation exercise.

@@

The LFM contribution to the AME comes in the multiplicative term: $\mathbf{u}_{i}^{\top} \mathbf{D} \mathbf{v}_{j}=\sum_{k \in K} d_{k} u_{ik} v_{jk}$. $K$ denotes the dimensions of the latent space. This model posits a latent vector of characteristics $\mathbf{u_{i}}$ and $\mathbf{v_{j}}$ for each sender $i$ and receiver $j$. The similarity or dissimilarity of these vectors will then influence the likelihood of activity, and provides a representation of third order interdependencies \citep{minhas:etal:2016:arxiv}. The LFM parameters are estimated by a process similar to computing the singular value decomposition (SVD) of the observed network. When taking the SVD we factorize our observed network into the product of three matrices: $\mathbf{U}, \mathbf{D}, \text{ and }, \mathbf{V}$. This provides us with a low-dimensional representation of our original network.\footnote{The dimensions of $\mathbf{U}$ and $\mathbf{V}$ are $n \times K$ and $\mathbf{D}$ is a $K \times K$ diagonal matrix.} Values in $\mathbf{U}$ provide a representation of how stochastically equivalent actors are as senders in a network or, for example, how similar actors are in terms of who they initiate conflict with. $\hat{\mathbf{u}}_{i} \approx \hat{\mathbf{u}}_{j}$ would indicate that actor $i$ and $j$ initiate battles with similar third actors. $\mathbf{V}$ provide a similar representation but from the perspective of how similar actors are as receivers. The values in $\mathbf{D}$, a diagonal matrix, represent levels of homophily in the network.\footnote{Unlike traditional SVD, in the latent factor model, the singular values are not restricted to be positive, thus allowing us to account for both the presence and absence of homophily.}

By integrating the SRRM and LFM into a Bayesian probit framework, we are able to account for the underlying structure in dyadic data that if left unestimated would complicate any inferences we might wish to draw for the exogenous parameters. Parameter estimation in the AME takes place within the context of a Gibbs sampler in which we iteratively sample from the posterior distribution of the full conditionals for each parameter. Specifically, given initial values of $\{\bm\beta, \mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}\}$, the algorithm proceeds as follows until convergence:

>>>>

For similar reasons one also needs to take into account the shared dependence between observations that share a common receiver. The bottom-left panel, illustrates that sender and receiver type dependencies can also blend together. Specifically, actors who are more likely to send ties in a network tend to also be more likely to receive them. As a result, the rows and columns in an adjacency matrix are often correlated. For example, consider that trade flows both from and to many wealthy, developed countries. The bottom-right panel, highlights a second-order dependence, specifically, reciprocity. This is a dependency occurring within dyads involving the same actors whereby values of $y_{ij}$ and $y_{ji}$ are correlated. The concept of reciprocity has deep roots in the study of relations between states \citep{richardson:1960,keohane:1989}.

For most relational data, however, dependencies do not simply manifest at the nodal or dyadic level. More often we find significant evidence of higher-order structures that result from dependencies between multiple groups of actors. These dependencies arise because there may be a or some set of latent attributes between actors that affects their probability of interacting with one another \citep{zinnes:1967,wasserman:faust:1994}. In Figure~\ref{fig:thirdDeps} we provide a visualization of a simulated relational dataset wherein the nodes designate actors and edges between the nodes indicate that an interaction between the two took place. To highlight third-order dependence patterns, nodes with similar latent attributes are colored similarly.

@@

The visualization illustrates that the actors belonging to the same group have a higher likelihood of having an interaction with each other, whereas interactions across groups are rarer. A prominent example of a network with this type of structure was found by \citet{adamic:glance:2005}, who visualized the ways in which right and left leaning political blogs linked to one another preceding the 2004 United States Presidential Election. \citeauthor{adamic:glance:2005} find that the degree of interaction between right and left leaning blogs was minimal, and that most blogs linked to others that were politically similar. This showcases the types of higher-order dependencies that can emerge in relational data. First, the fact that interactions were determined by a shared attribute, in this case political ideology, is an example of \textit{homophily}. Homophily explains the emergence of patterns such as transitivity (``a friend of a friend is a friend'') and balance (``an enemy of a friend is an enemy''), which also have a long history in international relations. The other major type of meso-scopic feature that emerges in relational data is community structure \citep{mucha:etal:2010}, which is often formalized through the concept of stochastic equivalence \citep{anderson:etal:1992}. This concept simply refers to the idea that groups of nodes that act similarly in the network are stochastically equivalent. In the example we have laid out above each of the left leaning blogs would be considered stochastically equivalent to one another.

The presence of these dependencies in relational data points to the fact that there is a complex structure underlying the dyadic events that we observe, and that accounting for this structure is necessary if we are to adequately represent the data generating process. Of course, if one can specify each of the nodal, dyadic, and triadic set of attributes that influence interactions then one can be assured that the conditional independence assumption underlying standard approaches will be satisfied. However, it is rarely the case that we can do this even for TSCS data, thus we more than often make modeling decisions to account for that structure. Failing to account for the underlying structure in either TSCS or dyadic data leads to a number of well-known challenges: a) biased estimates of the effect of independent variables, b) uncalibrated confidence intervals, and c) poor predictive performance. Further by ignoring these potential interdependencies, we often ignore substantively interesting features of the problem. The study of international relations is founded on the relations among actors. Why ignore the interdependencies that led to the study of IR in the first place?

@@

To account for the dependencies that are prevalent in dyadic data, we turn to the AME model. The AME approach can be used to conduct inference on cross-sectional and longitudinal networks with binary, ordinal, or continuous linkages. It is flexible and easy to use for analyzing the kind of relational data often found in social science. It accounts for nodal and dyadic dependence patterns, as well as higher-order dependencies such as homophily and stochastic equivalence. \citet{hoff:2015:arxiv,minhas:etal:2016:arxiv} provide a detailed introduction to this framework, and the latter piece also details how it contrasts with alternative network-based approaches. Here we provide a brief review and then move to a simulation exercise.

@@

The LFM contribution to the AME comes in the multiplicative term: $\mathbf{u}_{i}^{\top} \mathbf{D} \mathbf{v}_{j}=\sum_{k \in K} d_{k} u_{ik} v_{jk}$. $K$ denotes the dimensions of the latent space. This model posits a latent vector of characteristics $\mathbf{u_{i}}$ and $\mathbf{v_{j}}$ for each sender $i$ and receiver $j$. The similarity or dissimilarity of these vectors will then influence the likelihood of activity, and provides a representation of third-order interdependencies \citep{minhas:etal:2016:arxiv}. The LFM parameters are estimated by a process similar to computing the singular value decomposition (SVD) of the observed network. When taking the SVD we factorize our observed network into the product of three matrices: $\mathbf{U}, \mathbf{D}, \text{ and }, \mathbf{V}$. This provides us with a low-dimensional representation of our original network.\footnote{The dimensions of $\mathbf{U}$ and $\mathbf{V}$ are $n \times K$ and $\mathbf{D}$ is a $K \times K$ diagonal matrix.} Values in $\mathbf{U}$ provide a representation of how stochastically equivalent actors are as senders in a network or, for example, how similar actors are in terms of who they initiate conflict with. $\hat{\mathbf{u}}_{i} \approx \hat{\mathbf{u}}_{j}$ would indicate that actor $i$ and $j$ initiate battles with similar third actors. $\mathbf{V}$ provide a similar representation but from the perspective of how similar actors are as receivers. The values in $\mathbf{D}$, a diagonal matrix, represent levels of homophily in the network.\footnote{Unlike traditional SVD, in the latent factor model, the singular values are not restricted to be positive, thus allowing us to account for both the presence and absence of homophily.}

By integrating the SRRM and LFM into a Bayesian probit framework, we can account for the underlying structure in dyadic data that if left unestimated would complicate any inferences we might wish to draw for the exogenous parameters. Parameter estimation in the AME takes place within the context of a Gibbs sampler in which we iteratively sample from the posterior distribution of the full conditionals for each parameter. Specifically, given initial values of $\{\bm\beta, \mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}\}$, the algorithm proceeds as follows until convergence:

### 2017-10-29 paper/3_sim.tex
max, you are killing me.

We utilize a simulation study to highlight the utility of AME as an inferential tool for dyadic analysis. The goal of the simulation is to assess how well AME can provide unbiased and well calibrated estimates of coefficient parameters in the presence of unobserved dependencies. Specifically, we are concerned with conducting inference on regression parameters of a linear model for a network in the case where there is an omitted variable. Say that the true data-generating process for a particular $Y$ is given by:

@@

The first model corresponds to the ``naive'' approach in which little is explicitly done to account for latent dependencies in dyadic data. In the second model, we account for dependencies in dyadic data using the AME framework described in the previous section. For both the first and second models, we are simply estimating a linear model of $X$ on $Y$, and assessing the extent to which inference on the regression parameters are complicated in the presence of unobserved dependencies, $W$. In the last model, we provide an illustration of the ideal case in which we have observed and measured $W$ and are able to include it in our specification for $Y$. The oracle case provides an important benchmark for the naive and AME approaches.

@@

Next, we estimate the 95\% confidence interval for the three models in each of the simulations and estimate the proportion of times that the true value fell in those intervals. The results are summarized in Figure~\ref{fig:ameCalib}, and again we see that the AME approach performs as well as the oracle, while the naive approach performs poorly by comparison.

\begin{figure}
	\centering
	\caption{Proportion of times when true value fell in the estimated 95\% confidence interval for the naive, AME, and oracle models from 1,000 simulations.}
	\label{fig:ameCalib}
	\includegraphics[width=1\textwidth]{ameSimCover_all.pdf} \\
\end{figure}

Last, beyond obtaining less biased and better calibrated parameter estimates, a key benefit of the AME framework is that we can also estimate unobserved dependencies through the random effects structure of the model. In the case of the data generating process for $Y$, $W$ is set as an unobserved dyadic covariate that had a homophilous effect on $Y$. Homophilous because $W$ within this framework is simply an example of a dyadic attribute involving $i$ and $j$ that positively affects the degree to which they will interact with one another, i.e., $y_{ij}$. This type of unobserved dependency will be captured through the multiplicative effects portion of the model, $\mathbf{U}^{\top} \mathbf{D} \mathbf{V}$. To estimate how well the model performs in doing this we recover the multiplicative effects term for each simulation and calculate the correlation between it and the unobserved dependency, $W$.\footnote{Specifically, since both the multiplicative effects term and $W$ are continuous dyadic variables, we just calculate the Pearson correlation coefficient.} We visualize the distribution of the correlations from each of the 1,000 simulations in Figure~\ref{fig:ameCorr} for the case where the number of actors was set to 100 (top pair of panels) and 50. Additionally, we calculate the median across the correlations and display that result using a vertical line. For both $n=50$ and $n=100$, we find that the multiplicative effects perform very well in capturing the unobserved dependency, which indicates that the structure provided by this framework is not simply capturing noise but can be used as a tool to estimate unobserved structure.

>>>>

We utilize a simulation study to highlight the utility of AME as an inferential tool for dyadic analysis. The goal of the simulation is to assess how well AME can provide unbiased and well-calibrated estimates of coefficient parameters in the presence of unobserved dependencies. Specifically, we are concerned with conducting inference on regression parameters of a linear model for a network in the case where there is an omitted variable. Say that the true data-generating process for a particular $Y$ is given by:

@@

The first model corresponds to the ``naive'' approach in which little is explicitly done to account for latent dependencies in dyadic data. In the second model, we account for dependencies in dyadic data using the AME framework described in the previous section. For both the first and second models, we are simply estimating a linear model of $X$ on $Y$, and assessing the extent to which inference on the regression parameters are complicated in the presence of unobserved dependencies, $W$. In the last model, we provide an illustration of the ideal case in which we have observed and measured $W$ and include it in our specification for $Y$. The oracle case provides an important benchmark for the naive and AME approaches.

@@

Next, we estimate the 95\% confidence interval for the three models in each of the simulations and estimate the proportion of times that the true value fell within those intervals. The results are summarized in Figure~\ref{fig:ameCalib}, and again we see that the AME approach performs as well as the oracle, while the naive approach performs poorly by comparison.

\begin{figure}
	\centering
	\caption{Proportion of times the true value fell within the estimated 95\% confidence interval for the naive, AME, and oracle models from 1,000 simulations.}
	\label{fig:ameCalib}
	\includegraphics[width=1\textwidth]{ameSimCover_all.pdf} \\
\end{figure}

Last, beyond obtaining less biased and better-calibrated parameter estimates, a key benefit of the AME framework is that we can also estimate unobserved dependencies through the random effects structure of the model. In the case of the data generating process for $Y$, $W$ is set as an unobserved dyadic covariate that had a homophilous effect on $Y$. Homophilous because $W$ within this framework is simply an example of a dyadic attribute involving $i$ and $j$ that positively affects the degree to which they will interact with one another, i.e., $y_{ij}$. This type of unobserved dependency will be captured through the multiplicative effects portion of the model, $\mathbf{U}^{\top} \mathbf{D} \mathbf{V}$. To estimate how well the model performs in doing this we recover the multiplicative effects term for each simulation and calculate the correlation between it and the unobserved dependency, $W$.\footnote{Specifically, since both the multiplicative effects term and $W$ are continuous dyadic variables, we just calculate the Pearson correlation coefficient.} We visualize the distribution of the correlations from each of the 1,000 simulations in Figure~\ref{fig:ameCorr} for the case where the number of actors was set to 100 (top pair of panels) and 50. Additionally, we calculate the median across the correlations and display that result using a vertical line. For both $n=50$ and $n=100$, we find that the multiplicative effects perform very well in capturing the unobserved dependency, which indicates that the structure provided by this framework is not simply capturing noise but can be used as a tool to estimate unobserved structure.

### 2017-10-29 paper/4_app.tex
max, you are killing me.

\begin{table}[ht]
\centering
\caption{Here we provide a brief summary of the key variable in each of the five replications and a note about whether or not the finding is replicated when using our network based approach.}
	\begin{tabular}{l p{7cm} l} \toprule
		\multirow{2}{*}{Study} & \multirow{2}{*}{Central Finding} &  Replicates \\ 
			& &  in a Network Model? \\ \toprule
		Reiter \& Stam (2003) & Personalist Regimes Attack Democracies, Not Vice Versa & {Does Not Replicate} \\ \midrule
		McDonald (2004) & Lower Trade Barriers and Higher Trade Lead to Peace & {Does Not Replicate}\\ \midrule
		Rose (2004) & WTO Membership Does not Affect Trade & {Partially Replicates}\\ \midrule
		Weeks (2012) & Bosses, Juntas and Strongmen are more Aggressive, Machines are Not & {Does Not Replicate} \\\midrule
		Gibler (2017) & Power Parity at Time of Entry to International System Inceases Conflict & {Does Not Replicate}\\ \bottomrule
	\end{tabular}
	\label{tab:modelFindingSumm}
\end{table}

@@

\begin{table}[ht]
\centering
	\begin{tabular}{l|l cc}
	~ & ~ & GLM & AME \\
	\toprule
\multirow{2}{*}{Reiter \& Stam (2003)} & Area Under ROC Curve, AUC-ROC & 0.92 & {0.96} \\
		~ & Area Under PR Curve, AUC-PR & 0.08 &  {0.15} \\		\midrule
\multirow{2}{*}{McDonald (2004)} & AUC-ROC & 0.92 &  {0.99} \\
		~ & AUC-PR & 0.13 &  {0.28} \\		\midrule
\multirow{2}{*}{Rose (2004)} & RMSE & 3.23 &  {1.99} \\
		~ & RMDSE & 2.01 &  {1.06} \\	\midrule
\multirow{2}{*}{Weeks (2012)} & AUC-ROC & 0.64 &  {0.97} \\
		~ & AUC-PR & 0.00 &  {0.15} \\		\midrule
\multirow{2}{*}{Gibler (2017)} & AUC-ROC & 0.52 &  {0.91} \\
		~ & AUC-PR & 0.00 &  {0.08} \\			\bottomrule
	\end{tabular}
	\caption{Here we provide a summary of the out-of-sample performance based on our cross-validation strategy for each of the five replications when using the standard dyadic approach and our network based approach. Four of the five studies involved a binary dependent variable, so for those measures we provide area under the curve (AUC) statistics. The Rose study involved a Gaussian dependent variable and for that we use the root mean squared error (RMSE) and root median squared error (RMDSE). %Cases in which our network based approach outperformed the standard approach are highlighted in bold.
	}
	\label{tab:modelPerfSumm}
\end{table}
\FloatBarrier

@@

Our re-estimation with AME reveals that trade relations are highly interdependent and exhibit important patterns of transitivity.  Or, in other words, if countries $i$ and $j$ are highly dependent and countries $j$ and $k$ are also highly dependent, then we are likely to observe high dependency between countries $i$ and $k$. This indicates that conflict is less likely between members of a trade community. Once we control for these dependencies, we can more clearly interpret the positive link between trade and conflict.  The most striking thing is that AME finds a positive conditional association between trade dependence and conflict ($\hat{\beta}= 18.4, \quad \sigma_{\hat{\beta}} = 28.6$), while the comparable numbers for the logistic regression found in the original articles are negative ($\hat{\beta}= -22.2, \quad \sigma_{\hat{\beta}} = 15.2$). At the same time, the AME has ROC and PR curves (shown in the Appendix) that dominate the results found in McDonald (2004).
% Our findings support those of Traag and Lupu (2013) which argue that indirect trade relations reduce the probability of conflict.

@@

When we compare the results of Rose's original OLS model to the AME model accounting for network dependencies, the results are generally similar. The main result of the model---the null effect of membership in the WTO, as represented by the ``One-In" and ``Both-In" variables---does not prevail in the AME model. In the original, this negative linkage is not significant, but in the AME estimation this negative linkage is more precisely estimated. However, while in the original model there was a clear positive relationship between Real GDP and Trade, most of this effect vanishes in the AME model. The random effects shown in Figure~\ref{fig:roser} reveal the cause of much of this divergence. Here, the states with the most positive random effects are also states with high GDP, though not necessarily high GDP/capita.\footnote{Note: Qatar exhibits strongly negative random effects.} Thus, the effect of GDP in the original model was, in part, an artifact of first-order dependencies. Most of the other results of the model are constant across each model, though some geographic features, such as islands and landlocked states, have a more clear effect on trade once we account for these network dependencies. When we account for network interdependencies, we observe a markedly lower Root Mean Squared Error out-of-sample --- $3.23$ for the OLS model and $1.77$ for the AME model. While the AME model replicate the main result, there are substantial and substantive improvements gleaned from the AME results.

@@

The re-estimation of \citet{weeks:2012} has the sharpest divergence between the GLM results and those of the AME Model. In Weeks' initial models,  she finds that machines are less prone to initiate conflict than the reference category, whereas Juntas, Bosses and Strong-men are more conflict-prone. When we examine the AME results, we find that none of these values are distinguishable from zero. Similarly, we find less pronounced effects for military capabilities. One explanation for this divergence is the AME model's ability to account for third-order effects. Inspection of the multiplicative effects in Figure~\ref{fig:weekscirc} reveals a number of clusters of states which exhibit structural equivalence---in the top left corner we see the US, the UK, Germany, Canada and Israel. These states cluster together in the outer ring of this visualization because they tend to send conflicts to similar targets.\footnote{For details on how to interpret this type of visualization see \citet{minhas:etal:2016:arxiv}.} Conversely, in the bottom right of the outer ring, we see a cluster of authoritarian countries: Iraq, Russia, Syria, North Korea, and China. In the inner ring, the proximity of countries is determined by the degree to which they receive conflict from the same countries. In general, the clusters found on both the inner and outer rings have similar governmental types (Iraq, Syria, Libya, and North Korea all fell under the ``boss" category). In the GLM, which ignores these third-order dependencies, many of these results might have been attributed to regime type. Structural equivalence is present even when accounting for nodal characteristics like regime type.  The AME model, on the other hand, shows that this can be specified in terms of the interdependencies captured by the multiplicative effects.

>>>>

\begin{table}[ht]
\centering
\caption{Here we provide a brief summary of the key variable in each of the five replications and a note about whether or not the finding is replicated when using our network-based approach.}
	\begin{tabular}{l p{7cm} l} \toprule
		\multirow{2}{*}{Study} & \multirow{2}{*}{Central Finding} &  Replicates \\ 
			& &  in a Network Model? \\ \toprule
		Reiter \& Stam (2003) & Personalist Regimes Attack Democracies, Not Vice Versa & {Does Not Replicate} \\ \midrule
		McDonald (2004) & Lower Trade Barriers and Higher Trade Lead to Peace & {Does Not Replicate}\\ \midrule
		Rose (2004) & WTO Membership Does not Affect Trade & {Partially Replicates}\\ \midrule
		Weeks (2012) & Bosses, Juntas, and Strongmen are more Aggressive, Machines are Not & {Does Not Replicate} \\\midrule
		Gibler (2017) & Power Parity at Time of Entry to International System Inceases Conflict & {Does Not Replicate}\\ \bottomrule
	\end{tabular}
	\label{tab:modelFindingSumm}
\end{table}

@@

\begin{table}[ht]
\centering
	\begin{tabular}{l|l cc}
	~ & ~ & GLM & AME \\
	\toprule
\multirow{2}{*}{Reiter \& Stam (2003)} & Area Under ROC Curve, AUC-ROC & 0.92 & {0.96} \\
		~ & Area Under PR Curve, AUC-PR & 0.08 &  {0.15} \\		\midrule
\multirow{2}{*}{McDonald (2004)} & AUC-ROC & 0.92 &  {0.99} \\
		~ & AUC-PR & 0.13 &  {0.28} \\		\midrule
\multirow{2}{*}{Rose (2004)} & RMSE & 3.23 &  {1.99} \\
		~ & RMDSE & 2.01 &  {1.06} \\	\midrule
\multirow{2}{*}{Weeks (2012)} & AUC-ROC & 0.64 &  {0.97} \\
		~ & AUC-PR & 0.00 &  {0.15} \\		\midrule
\multirow{2}{*}{Gibler (2017)} & AUC-ROC & 0.52 &  {0.91} \\
		~ & AUC-PR & 0.00 &  {0.08} \\			\bottomrule
	\end{tabular}
	\caption{Here we provide a summary of the out-of-sample performance based on our cross-validation strategy for each of the five replications when using the standard dyadic approach and our network-based approach. Four of the five studies involved a binary dependent variable; for those measures, area under the curve (AUC) statistics are reported. The Rose study involved a Gaussian dependent variable and for that we use the root mean squared error (RMSE) and root median squared error (RMDSE). 
	}
	\label{tab:modelPerfSumm}
\end{table}
\FloatBarrier

@@

Our re-estimation with AME reveals that trade relations are highly interdependent and exhibit important patterns of transitivity.  Or, in other words, if countries $i$ and $j$ are highly dependent and countries $j$ and $k$ are also highly dependent, then we are likely to observe high dependency between countries $i$ and $k$. This indicates that conflict is less likely among members of a trade community. Once we control for these dependencies, we can more clearly interpret the positive link between trade and conflict.  The most striking thing is that AME finds a positive conditional association between trade dependence and conflict ($\hat{\beta}= 18.4, \quad \sigma_{\hat{\beta}} = 28.6$), while the comparable numbers for the logistic regression found in the original articles are negative ($\hat{\beta}= -22.2, \quad \sigma_{\hat{\beta}} = 15.2$). At the same time, the AME has ROC and PR curves (shown in the Appendix) that dominate the results found in McDonald (2004).
% Our findings support those of Traag and Lupu (2013) which argue that indirect trade relations reduce the probability of conflict.

@@

When we compare the results of Rose's original OLS model to the AME model accounting for network dependencies, the results are similar. The main result of the model---the null effect of membership in the WTO, as represented by the ``One-In" and ``Both-In" variables---does not prevail in the AME model. In the original, this negative linkage is not significant, but in the AME estimation this negative linkage is more precisely estimated. However, while in the original model there was a clear positive relationship between Real GDP and Trade, most of this effect vanishes in the AME model. The random effects shown in Figure~\ref{fig:roser} reveal the cause of much of this divergence. Here, the states with the most positive random effects are also states with high GDP, though not necessarily high GDP/capita.\footnote{Note: Qatar exhibits strongly negative random effects.} Thus, the effect of GDP in the original model was, in part, an artifact of first-order dependencies. Most of the other results of the model are constant across each model, though some geographic features, such as islands and landlocked states, have a more clear effect on trade once we account for these network dependencies. When we account for network interdependencies, we observe a markedly lower Root Mean Squared Error out-of-sample --- $3.23$ for the OLS model and $1.77$ for the AME model. While the AME model replicate the main result, there are substantial and substantive improvements gleaned from the AME results.

@@

The re-estimation of \citet{weeks:2012} has the sharpest divergence between the GLM results and those of the AME Model. In Weeks' initial models,  she finds that machines are less prone to initiate conflict than the reference category, whereas Juntas, Bosses, and Strong-men are more conflict-prone. When we examine the AME results, we find that none of these values are distinguishable from zero. Similarly, we find less pronounced effects for military capabilities. One explanation for this divergence is the AME model's ability to account for third-order effects. Inspection of the multiplicative effects in Figure~\ref{fig:weekscirc} reveals a number of clusters of states which exhibit structural equivalence---in the top left corner we see the US, the UK, Germany, Canada, and Israel. These states cluster together in the outer ring of this visualization because they tend to send conflicts to similar targets.\footnote{For details on how to interpret this type of visualization see \citet{minhas:etal:2016:arxiv}.} Conversely, in the bottom right of the outer ring, we see a cluster of authoritarian countries: Iraq, Russia, Syria, North Korea, and China. In the inner ring, the proximity of countries is determined by the degree to which they receive conflict from the same countries. In general, the clusters found on both the inner and outer rings have similar governmental types (Iraq, Syria, Libya, and North Korea all fell under the ``boss" category). In the GLM, which ignores these third-order dependencies, many of these results might have been attributed to regime type. Structural equivalence is present even when accounting for nodal characteristics like regime type.  The AME model, on the other hand, shows that this can be specified in terms of the interdependencies captured by the multiplicative effects.

### 2017-10-29 paper/5_concl.tex
max, you are killing me.

To explore this approach in the context of international relations we have presented two broad analyses. The first is a simulation where the characteristics of the network are known. This shows that, when there are unobserved dependencies, the AME approach is less biased in terms of parameter estimation compared with standard approach employed in international relations to study dyadic data (i.e., GLM models). The second is a replication of five prominent studies that have been published recently using a broad range of dyadic data to draw inferences about international relations.  These five studies have been replicated with the original research designs, each of which used a statistical method that assumes the dyadic data are all independent from one another.  We then re-analyzed each study using the AME model.  In every case, we found that the AME approach provided a) increased precision of estimation, b) better out-of-sample fits, c) evidence of 1st-, 2nd-, and 3rd-order dependencies that were overlooked in the original studies.\footnote{The Appendix contains performance data on all of these replications, as well as sample code illustrating how to undertake AME analysis using \texttt{amen}.} In several cases, the new approach overturns the basic findings of the original research.  This leads us to speculate that many of the findings in the international relations literature may be fragile in the sense that they can only be obtained under stringent assumptions that can not possibly be valid.  This in turn leads to a certain arbitrariness in some research findings, which might lead to puzzles that are more apparent than real \citep{zinnes:1980}.

>>>>

To explore this approach in the context of international relations we have presented two broad analyses. The first is a simulation where the characteristics of the network are known. This shows that, when there are unobserved dependencies, the AME approach is less biased in terms of parameter estimation compared with standard approach employed in international relations to study dyadic data (i.e., GLM models). The second is a replication of five prominent studies that have been published recently using a broad range of dyadic data to draw inferences about international relations.  These five studies have been replicated with the original research designs, each of which used a statistical method that assumes the dyadic data are all independent from one another.  We then re-analyzed each study using the AME model.  In every case, we found that the AME approach provided a) increased precision of estimation, b) better out-of-sample fits, c) evidence of 1st-, 2nd-, and 3rd-order dependencies that were overlooked in the original studies.\footnote{The Appendix contains performance data on all of these replications, as well as sample code illustrating how to undertake AME analysis using \texttt{amen}.} In several cases, the new approach overturns the basic findings of the original research.  This leads us to speculate that many of the findings in the international relations literature may be fragile in the sense that they can only be obtained under stringent assumptions that cannot possibly be valid.  In turn, this leads to a certain arbitrariness in some research findings, which might lead to puzzles that are more apparent than real \citep{zinnes:1980}.
